Learning Features and Abstract Actions for Computing Generalized Plans

### 2018-11-14 doc/camera-ready/paper.tex
Last minute changes

Bonet and Geffner's generalized planning formulation assumes that the
features and abstract actions are given. In this work, we address this limitation
by  \emph{showing  how the features and abstract actions  can be learned from
the primitive predicates used to define the instances and from sampled state transitions.}
For example,  the general policy for achieving $clear(x)$ % of \citeay{bonet:ijcai2018}
is obtained using a FOND planner on an abstraction that consists of  a  boolean
feature $H$,  that tracks whether the gripper holds a block,  a numerical feature
$n(x)$ that counts the number of blocks above $x$, and two abstract actions:
one with preconditions $\neg H$ and $n(x) > 0$, and effects $H$ and $n(x)\mminus$
(decrement of $n(x)$), and the other with precondition $H$ and effect $\neg H$.
We here show how to obtain such policies from STRIPS instances alone,
without having to provide the features and the abstract actions by hand.

@@

The paper is organized as follows. We first provide background and then show
how abstract actions can be learned by enforcing soundness and
completeness over a set of samples from a pool of candidate features derived
from the domain predicates. The computational model is then summarized,
followed by experimental results and a discussion.

@@

A \emph{feature} $f$ for a class $\Q$ of problems represents a function $\phi_f$
that takes an   instance $P$ from $\Q$ and a state $s$ reachable in $P$, and results
in a value denoted as $\phi_f(s)$.
A feature is \emph{boolean} if it results in a boolean value, and \emph{numeric}
if it results in a numerical value, that we assume to be a non-negative integer.
For example, $H$ and $n(x)$ represent two features in the Blocksworld:
the first is a boolean feature that tracks whether the gripper is empty, while
the second is numerical and tracks the number of blocks above $x$.

@@

An action $a$ and an abstract action $\bar{a}=\abst{Pre}{\Eff}$  \emph{have  the same qualitative effects over the features $F$
in a state $s$} when both are applicable in $s$ and have the same effects on the boolean features and
the same \emph{qualitative effects} on the numerical features. If $s'$ is the result of applying $a$ in $s$,
this means \cite{bonet:ijcai2018}:
1)~$p$ becomes true in $s'$,
i.e., $\phi_p(s')$ is true and $\phi_p(s)$  is false,
iff $p \in \Eff$ and $\neg p \in Pre$, 2)~$p$ becomes false in $s'$, i.e., $\phi_p(s')$ is false and $\phi_p(s)$  is true,
iff $\neg p \in \Eff$ and $p \in Pre$, 3)~$n$ increases in $s'$, i.e., $\phi_n(s') > \phi_n(s)$, iff $n\pplus \in \Eff$,
and 4)~$n$ decreases  in $s'$, i.e., $\phi_n(s') < \phi_n(s)$, iff $n\mminus \in \Eff$.

@@

Soundness and completeness are the key  properties that  enable us to reason about the collection of
instances $\Q$ in terms of abstract actions that are defined at the level of the features:

@@

\medskip
\begin{example}
  The abstract action set $A_F = \{\bar{a}, \bar{a}'\}$ where $\bar{a} = \abst{\neg H,n(x)>0}{H,n(x)\mminus}$
  and $\bar{a}'=\abst{H}{\neg H}$, defined over the above feature set $F=\{H,n(x)\}$,
  is sound but not complete relative to $\Q_{clear}$.
  The actions in $A_F$ capture the concrete actions that pick blocks from above $x$ and put them away,
  but not other actions like stacking a block above $x$, or picking a block that is not above $x$.
\end{example}

@@

Different assignments $\sigma$ result in different feature sets $F_{\sigma}$
and abstract action sets $A_{\sigma}$. Simpler and more meaningful features can be found
by minimizing a cost measure such as $|F_{\sigma}|$
or, more generally, $\sum_{f \in F_{\sigma}} cost(f)$.
We achieve this by casting our problem
into a constrained optimization problem, more specifically, a Weighted Max-SAT problem,
where clauses resulting from the theories $T$ or $T_G$ are taken as \emph{hard} clauses,
and we add, for each feature $f \in \F$, a
\emph{soft} unit clause $\neg selected(f)$ with weight $cost(f)$.
The cost measure that is used in the experiments is defined below.

@@

Finally, the formulation can be extended to handle non-deterministic actions.
For this, it is sufficient to replace the formula \eqref{eq:bridge1}
that links $D_1$ and $D_2$ atoms  with a formula asserting that,
if two states $s$ and $t$ are indistinguishable (i.e., $\neg D_1(s,t)$), then
for each action $a$ in $s$, there is an action $b$ in  $t$
such that the set of transitions $(t,b,t')$ generated by $b$ in $t$
cannot be distinguished from the set of transitions $(s,a,s')$ generated by $a$ in $s$.
For this, transitions $(s,s')$ in the sample would need to be tagged with the actions
that generated them as  $(s,a,s')$.

>>>>

Bonet and Geffner's formulation of generalized planning  assumes that the
features and abstract actions are given. In this work, we address this limitation
by  \emph{showing  how the features and abstract actions  can be learned from
the primitive predicates used to define the instances and from sampled state transitions.}
For example,  the general policy for achieving $clear(x)$ % of \citeay{bonet:ijcai2018}
is obtained using a FOND planner on an abstraction that consists of  a  boolean
feature $H$,  that tracks whether the gripper holds a block,  a numerical feature
$n(x)$ that counts the number of blocks above $x$, and two abstract actions:
one with preconditions $\neg H$ and $n(x) > 0$, and effects $H$ and $n(x)\mminus$
(decrement of $n(x)$), and the other with precondition $H$ and effect $\neg H$.
We here show how to obtain such policies from STRIPS instances alone,
without having to provide the features and the abstract actions by hand.

@@

The paper is organized as follows. We first provide the relevant background, and then show
how features and abstract actions can be learned by enforcing soundness and
completeness over a set of samples, from a pool of candidate features derived
from the domain predicates. The computational model is then summarized,
followed by experimental results and a discussion.

@@

A \emph{feature} $f$ for a class $\Q$ of problems represents a function $\phi_f$
that takes an instance $P$ from $\Q$ and a state $s$ reachable in $P$, and results
in a value $\phi_f(s)$.
A feature is \emph{boolean} if it results in boolean values, and \emph{numeric}
if it results in numerical values. %, that we assume to be a non-negative integer.
For example, $H$ and $n(x)$ represent two features in the Blocksworld:
the first is a boolean feature that tracks whether the gripper is empty, while
the second is numerical and tracks the number of blocks above $x$.

@@

An action $a$ and an abstract action $\bar{a}=\abst{Pre}{\Eff}$  \emph{have  the same qualitative effects over the features $F$
in a state $s$} when both are applicable in $s$ with the same effects on the boolean features and
the same \emph{qualitative effects} on the numerical features. If $s'$ is the result of applying $a$ in $s$,
this means \cite{bonet:ijcai2018}:
1)~$p \in \Eff$ and $\neg p \in Pre$ iff
$\phi_p(s)$ is false and $\phi_p(s')$ is true (i.e., $p$ becomes true),
%
2)~$\neg p \in \Eff$ and $p \in Pre$ iff
$\phi_p(s)$ is true and $\phi_p(s')$ is false ($p$ becomes false),
%
3)~$n\pplus \in \Eff$ iff $\phi_n(s') > \phi_n(s)$ ($n$ increases),
and 
%
4)~$n\mminus \in \Eff$ iff $\phi_n(s') < \phi_n(s)$ ($n$ decreases).

@@

Soundness and completeness are the key  properties that  enable us to reason about the collection of
instances $\Q$ in terms of abstract actions that operate at the level of the features:

@@

\medskip
\begin{example}
  The abstract action set $A_F = \{\bar{a}, \bar{a}'\}$ where $\bar{a} = \abst{\neg H,n(x)>0}{H,n(x)\mminus}$
  and $\bar{a}'=\abst{H}{\neg H}$, defined over the feature set $F=\{H,n(x)\}$,
  is sound but not complete relative to $\Q_{clear}$.
  The actions in $A_F$ capture the concrete actions that pick blocks from above $x$ and put them away,
  but not other actions like stacking a block above $x$, or picking a block that is not above $x$.
\end{example}

@@

Different assignments $\sigma$ result in different feature sets $F_{\sigma}$
and abstract action sets $A_{\sigma}$. Simpler and more meaningful features can be found
by minimizing a cost measure such as $|F_{\sigma}|$
or, more generally, $\sum_{f \in F_{\sigma}} cost(f)$.
We achieve this by casting our problem
into a constrained optimization problem, more specifically, a Weighted Max-SAT problem,
where clauses resulting from the theories $T$ or $T_G$ are taken as \emph{hard} clauses,
and we add, for each feature $f \in \F$, a
\emph{soft} unit clause $\neg selected(f)$ with weight $cost(f)$, defined below.

@@

Finally, our formulation can be extended to handle non-deterministic actions.
For this, it is sufficient to replace the formula \eqref{eq:bridge1}
that links $D_1$ and $D_2$ atoms  with a formula asserting that,
if two states $s$ and $t$ are indistinguishable (i.e., $\neg D_1(s,t)$), then
for each action $a$ in $s$, there is an action $b$ in  $t$
such that the set of transitions $(t,b,t')$ generated by $b$ in $t$
cannot be distinguished from the set of transitions $(s,a,s')$ generated by $a$ in $s$.
For this, transitions $(s,s')$ in the sample would need to be tagged with the actions
that generated them as  $(s,a,s')$.
